---
title: "ISL_chapter_5_lab"
format: html
editor: visual
---

# ISL Chapter 5 Lab

## Libraries

```{r}
library(tidyverse)
library(ISLR2)
library(boot)
```

## The Validation Set Approach

First, we set the seed and select a subset of half of the original 392 observations.

```{r}
set.seed(1)
training <- sample(392, 196)
```

We then fit a linear regression model to predict MPG with horsepower using the raining data. We then calculate its estimated MSE.

```{r}
lm.fit <- lm(mpg ~ horsepower, data = Auto, subset = training)
attach(Auto)
mean((mpg - predict(lm.fit, Auto))[-training] ^ 2)
```

Next, we fit quadratic and cubic regression models.

```{r}
qm.fit <- lm(mpg ~ poly(horsepower, 2), data = Auto, subset = training)
mean((mpg - predict(qm.fit, Auto))[-training] ^ 2)

cm.fit <- lm(mpg ~ poly(horsepower, 3), data = Auto, subset = training)
mean((mpg - predict(cm.fit, Auto))[-training] ^ 2)
```

We can repeat this process with a different seed to see new results.

```{r}
set.seed(2)
training <- sample(392, 196)

lm.fit <- lm(mpg ~ horsepower, data = Auto, subset = training)
mean((mpg - predict(lm.fit, Auto))[-training]^2)

qm.fit <- lm(mpg ~ poly(horsepower, 2), data = Auto, subset = training)
mean((mpg - predict(qm.fit, Auto))[-training]^2)

cm.fit <- lm(mpg ~ poly(horsepower, 3), data = Auto, subset = training)
mean((mpg - predict(cm.fit, Auto))[-training]^2)
```

The trend with a seed value of 2, however, shows a similar result to a seed value of 1. The quadratic model seems to be the best-fitting model compared to a linear or cubic model.

## Leave-One-Out CV

The glm() function can perform linear regression without a family argument.

```{r}
glm.fit <- glm(mpg ~ horsepower, data = Auto)
cv.err <- cv.glm(Auto, glm.fit)
cv.err$delta
```

The delta vector contains the cross-validation results.

Changing the degree of the polynomial does not yield a significant drop after increasing from linear to quadratic.

```{r}
cv.error <- rep(0, 10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error[i] <- cv.glm(Auto, glm.fit)$delta[1]
}
cv.error
```

## k-Fold Cross-Validation

cv.glm() can be used to implement k-fold CV.

```{r}
set.seed(17)
cv.error.10 <- rep(0, 10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error.10[i] <- cv.glm(Auto, glm.fit, K = 10)$delta[1]
}
cv.error.10
```

As above, using a cubic or greater degree polynomial doesn't seem to yield a lower test error.
