---
title: "chapter_6_lab"
format: html
editor: visual
---\
---

# Libraries

Load the libraries.

```{r}
library(tidyverse)
library(glmnet)
library(ISLR2)
library(leaps)
```

# EDA and Preprocessing

This is some minor exploratory data analysis.

```{r}
names(Hitters)
dim(Hitters)
sum(is.na(Hitters$Salary))

Hitters <- na.omit(Hitters)
dim(Hitters)
```

I skipped the best subset selection portion of the lab: the rest of 6.5.1. This first section generates a matrix, transforming all qualitative predictors into dummy variables. I also fit the ridge regression model, with 100 different lambda values, spanning from 10\^-2 to 10\^10.

```{r}
X <- model.matrix(Salary ~ ., Hitters)[, -1]
y <- Hitters$Salary

grid <- 10^seq(10, -2, length = 100)
ridge.mod <- glmnet(X, y, alpha = 0, lambda = grid)
```

This cell will split the data into training and testing data.

```{r}
set.seed(1)
train <- sample(1:nrow(X), nrow(X) / 2)
test <- (-train)

y.test <- y[test]
```

# Ridge regression

This identifies different values of lambda, observes the coefficients, and gets the L2 norm, or the length by Euclidean distance.

```{r}
ridge.mod$lambda[5]
coef(ridge.mod)[, 5]
sqrt(sum(coef(ridge.mod)[-1, 5]^2))
```

```{r}
ridge.mod$lambda[25]
coef(ridge.mod)[, 25]
sqrt(sum(coef(ridge.mod)[-1, 25]^2))
```

```{r}
ridge.mod$lambda[95]
coef(ridge.mod)[, 95]
sqrt(sum(coef(ridge.mod)[-1, 95]^2))
```

The predict() function is used to get the ridge coefficients for a new lambda value.

```{r}
predict(ridge.mod, s = 100, type = "coefficients")[1:20, ]
```

We can fit the ridge regression model to the training set and gather its mean squared error (MSE) on the test set. This uses a lambda value of 4. The lower part of the cell just uses an intercept.

```{r}
ridge.mod <- glmnet(X[train, ], y[train], alpha = 0, lambda = grid, thresh = 1e-12)
ridge.pred <- predict(ridge.mod, s = 4, newx = X[test, ])
mean((ridge.pred - y.test) ^ 2)

mean((mean(y[train]) - y.test) ^ 2)
```

Here is a least squares model, or a ridge model with a lambda value of 0.

```{r}
ridge.pred <- predict(ridge.mod, s = 0, newx = X[test, ], exact = T, x = X[train, ], y = y[train])
mean((ridge.pred - y.test) ^ 2)

lm(y ~ X, subset = train)
predict(ridge.mod, s = 0, exact = T, type = "coefficients", x = X[train, ], y = y[train])[1:20, ]
```

Using the cross-validation function, I can find the optimal value of lambda.

```{r}
set.seed(1)
cv.out <- cv.glmnet(X[train, ], y[train], alpha = 0)
plot(cv.out)

best_lambda <- cv.out$lambda.min
best_lambda

ridge.pred <- predict(ridge.mod, s = best_lambda, newx = X[test, ])
mean((ridge.pred - y.test) ^ 2)
```

This refits the ridge model on the full data set.

```{r}
out <- glmnet(X, y, alpha = 0)
predict(out, type = "coefficients", s = best_lambda)[1:20, ]
```

This shows that a properly cross-validated ridge regression model can outperform a least squares model.

# Lasso regression

This cell fits a lasso model where alpha is equal to 1.

```{r}
lasso.mod <- glmnet(X[train, ], y[train], alpha = 1, lambda = grid)
plot(lasso.mod)
```

This cell identifies the best lambda value of \~9.29 and calculates is test set MSE of \~143,674.

```{r}
set.seed(1)
cv.out <- cv.glmnet(X[train, ], y[train], alpha = 1)
plot(cv.out)

best_lambda <- cv.out$lambda.min
best_lambda
lasso.pred <- predict(lasso.mod, s = best_lambda, newx = X[test, ])
mean((lasso.pred - y.test) ^ 2)
```

This cell fits a lasso regression model using the above best lambda value. It is worth noting that, by comparison to the ridge regression model, some of these coefficients are collapsed to zero.

```{r}
out <- glmnet(X, y, alpha = 1, lambda = grid)
lasso.coef <- predict(out, type = "coefficients", s = best_lambda)[1:20, ]
lasso.coef
```
