---
title: "figure 2.4"
format: html
editor: visual
---

# Creating Figure 2.4

## Load data

```{r}
library(tidyverse)
library(mvtnorm)
library(caret)
```

Create centers that are either orange or blue with std. dev. of 1; orange centers are centered on (0,1) and blue centers on (1,0), where the x and y values are independent of each other.

```{r}
orange_centers <- tibble(x = rnorm(10, 0, 1), y = rnorm(10, 1, 1))
blue_centers <- tibble(x = rnorm(10, 1, 1), y = rnorm(10, 0, 1))
```

```{r}
orange_matrix <- matrix(, nrow = 100, ncol = 3)
blue_matrix <- matrix(,nrow = 100, ncol = 3)

for (i in 1:100) {
  rand_blue <- sample(1:10)
  rand_orange <- sample(1:10)
  
  blue_center <- c(blue_centers$x[rand_blue], blue_centers$y[rand_blue])
  orange_center <- c(orange_centers$x[rand_orange], orange_centers$y[rand_orange])
  
  blue_point <- rnorm(2, mean = blue_center, sd = sqrt(1/5))
  orange_point <- rnorm(2, mean = orange_center, sd = sqrt(1/5))
  
  blue_matrix[i, ] <- c(blue_point, 0)
  orange_matrix[i, ] <- c(orange_point, 1)
}

colnames(blue_matrix) <- c("x", "y", "color")
colnames(orange_matrix) <- c("x", "y", "color")

blue <- as_tibble(blue_matrix)
orange <- as_tibble(orange_matrix)

training <- rbind(blue, orange)
```

```{r}
ggplot(data = training) +
  geom_point(mapping = aes(x = x, y = y, color = factor(color))) +
  scale_color_manual(values = c("0" = "blue", "1" = "orange"))
```

## Creating the figure

Figure 2.4 shows the training and testing error for various values of k in a K-nearest neighbors classifier, with a test set of size 10,000 and a training set as generated above. In addition, it shows the optimal Bayes error rate.

```{r}

```
